\batcave: Bayesian Analysis Tools for Context-Aware Variant Evaluation

### 2019-09-10 NAR_DRAFT.tex
note to fix figure 2
remove figure note
incredible struggle to get numeric and inline. back to author-year for now. looks to be OK for submission

\documentclass[a4,center,fleqn]{NAR}

\usepackage{NAR-natbib}
\bibliographystyle{NAR-natbib}
\usepackage{color}
\usepackage{xfrac}
\newcommand{\rngcomment}[1]{{\color{red}RNG: #1}}
\newcommand{\bkmcomment}[1]{{\color{blue}BKM: #1}}
\newcommand{\batcave}{BATCAVE }
\newcommand{\beginsupplement}{%
        \clearpage
        \onecolumn
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }
% Enter dates of publication
\copyrightyear{2008}
\pubdate{31 July 2009}
\pubyear{2009}
\jvolume{37}
\jissue{12}

@@

Methods for identifying true somatic mutations - i.e. variant calling -  from NGS data are an active area of research in bioinformatics.
The earliest widely used somatic variant callers aimed specifically at tumors, Mutect1 and Varscan2, used a combination of heuristic filtering and a model of sequencing errors to identify and score potential variants, setting a threshold for that score designed to balance sensitivity and specificity \citep{Koboldt2012,Cibulskis2013}.
Subsequent research gave rise to a number of alternate variant calling strategies including haplotype based callers \citep{Garrison2012},
joint genotype analysis (SomaticSniper, JointSNVMix2, Seurat, and CaVEMan,MuClone)\citep{Larson2012,Roth2012a,Christoforides2013,Jones2016,Dorri2019}, allele frequency based analysis (Strelka, MuTect, LoFreq, EBCall, deepSNV, LoLoPicker, and MuSE)\citep{Saunders2012,Wilm2012,Shiraishi2013b,Gerstung2012,Carrot-Zhang2017,Fan2016}, and a mixture of ensemble and deep learning methods (MutationSeq, SomaticSeq, SNooPer, and BAYSIC).
All of these methods have varying levels of complexity, and some are focused on specific types of data.
The one thing they all have in common is that they either implicitly or explicitly assume that the probability of a mutation occuring at a particular site is proportional to the overall mutation rate, and the same at every site in the genome.

@@

  \begin{figure*}
    \begin{center}
    \includegraphics{figures/fig_wes.pdf}
    \end{center}
    \caption{Model performance.
    \textbf{(a)} Precision-recall curves and \textbf{(b)} Receiver operating characteristic curves for 3 mutation signatures.
    Distribution of estimated true positive probabilities for true positive (top) and true negative (bottom) variants for \textbf{(c)} the MuTect model and \textbf{(d))} the \batcave model.
    A perfectly calibrated model would generate the diagonal line. Precision-recall curves for real data \textbf{(e)} and \textbf{(f)}}
  \label{NAR-wes_fig}
  \end{figure*}

>>>>

\documentclass[a4,center,fleqn]{NAR}
\bibliographystyle{plainnat}
\usepackage{color}
\usepackage{xfrac}

\newcommand{\rngcomment}[1]{{\color{red}RNG: #1}}
\newcommand{\bkmcomment}[1]{{\color{blue}BKM: #1}}
\newcommand{\batcave}{BATCAVE }
\newcommand{\beginsupplement}{%
        \clearpage
        \onecolumn
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }
% Enter dates of publication
\copyrightyear{2008}
\pubdate{31 July 2009}
\pubyear{2009}
\jvolume{37}
\jissue{12}

@@

Methods for identifying true somatic mutations - i.e. variant calling -  from NGS data are an active area of research in bioinformatics.
The earliest widely used somatic variant callers aimed specifically at tumors, Mutect1 and Varscan2, used a combination of heuristic filtering and a model of sequencing errors to identify and score potential variants, setting a threshold for that score designed to balance sensitivity and specificity \citep{Koboldt2012,Cibulskis2013}.
Subsequent research gave rise to a number of alternate variant calling strategies including haplotype based callers \citep{Garrison2012},
joint genotype analysis (SomaticSniper, JointSNVMix2, Seurat, and CaVEMan,MuClone)\cite{Larson2012,Roth2012a,Christoforides2013,Jones2016,Dorri2019}, allele frequency based analysis (Strelka, MuTect, LoFreq, EBCall, deepSNV, LoLoPicker, and MuSE)\citep{Saunders2012,Wilm2012,Shiraishi2013b,Gerstung2012,Carrot-Zhang2017,Fan2016}, and a mixture of ensemble and deep learning methods (MutationSeq, SomaticSeq, SNooPer, and BAYSIC).
All of these methods have varying levels of complexity, and some are focused on specific types of data.
The one thing they all have in common is that they either implicitly or explicitly assume that the probability of a mutation occuring at a particular site is proportional to the overall mutation rate, and the same at every site in the genome.

@@

  \begin{figure*}
    \begin{center}
    \includegraphics{figures/fig_wes.pdf}
    \end{center}
    \caption{Model performance.
    \textbf{(a)} Precision-recall curves and \textbf{(b)} Receiver operating characteristic curves for 3 mutation signatures.
    Distribution of estimated true positive probabilities for true positive (top) and true negative (bottom) variants for \textbf{(c)} the MuTect model and \textbf{(d))} the \batcave model.
    A perfectly calibrated model would generate the diagonal line. Precision-recall curves for real data \textbf{(e)} and \textbf{(f)}.}
  \label{NAR-wes_fig}
  \end{figure*}
